---
title: "Exploring the role of process model parameters in phantom interactions on RTs"
author: "Adriana F. Chávez De la Peña and Joachim Vandekerckhove"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    cards: false
    highlight: tango
    fig_width: 12 
    fig_height: 8 
    css: src/custom.css
---

```{r, echo = FALSE, message = FALSE}
library(here)
library(knitr)
library(kableExtra) 
library(tidyverse)
library(R2jags)

# Load custom plotting functions
source(here("src", "sample_summary-stats.R"))
source(here("src", "plot_verticalHist.R"))
source(here("src", "plot_posteriors.R"))
source(here("src", "get_Rhat.R"))

seed <- 123
```


# Introduction

Phantom interactions are scale-dependent statistically significant noncrossover interactions between two or more factors that disappear after appropriate scale transformation.

**Key characteristics:**

- **Scale-Dependent:** The interaction's presence or significance depends heavily on the scale used for the dependent variable.
- **Visual Cue:** On an interaction plot using the original scale, the lines representing the levels of one factor across the levels of another factor will be non-parallel but typically won't cross within the observed data range (hence "noncrossover"). After transformation, these lines often become (approximately) parallel.
- **Removable:** The interaction can often be "removed" by changing the scale, potentially simplifying the model interpretation to only main effects on the transformed scale.

Phantom interactions are an artifact of the original dependent variable measurement scale. They can't be used to conclude that the effect of one factor fundamentally changes depending on the level of another. As such, they highlight the importance of considering the appropriate scale for analyzing data, especially when effects might combine multiplicatively rather than purely additively.

# Test 1: <br>Single participant tested over multiple days and conditions

```{r}
# Simulation design settings:
J <- 2    # Design factors
L <- 2    # Levels per factor
K <- J*L  # Conditions
D <- 8    # Repeated observations (experimental sessions)
N <- 120  # Trials per condition
```

We generate data for a single participant in a binary decision task with $K =$ `r K` conditions defined by a simple `r L`-by-`r L` factorial design:

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create the data frame
my_data <- data.frame(
  FactorB = c("FactorB1", "FactorB2"),
  FactorA1 = c("Cond 1", "Cond 3"),
  FactorA2 = c("Cond 2", "Cond 4")
)

# Generate the table using kable
kable(my_data, 
      col.names = c("", "FactorA1", "FactorA2"), # Set column names, empty for the first
      align = 'lcc') %>%                        # Align columns (left, center, center)
  kable_styling(bootstrap_options = c("striped", "bordered"), # Add styling
                full_width = FALSE)             # Prevent table from spanning full page width
```


For each condition $k$, we generate summary statistics for the choice and response time data using the EZ drift diffusion model (EZ-DDM). As such, the accuracy rate $\dot{A}_k$, mean response time $\dot{MRT}_k$, and response time variance $\dot{VRT}_k$ are modeled as:

$$\left\{\dot{A}_k, \dot{MRT}_k, \dot{VRT}_k\right\} \sim \mbox{EZDDM}\left(\nu_{k}, \alpha_{k}, \tau\right )$$

where $\nu$, $\alpha$, and $\tau$ are the parameters of the simple drift diffusion model (the drift rate, boundary separation, and non-decision time, respectively).

We generate $D =$ `r D` summary statistics per condition, as if the participant completed the task on $D$ experimental sessions.

## Simulate data

### Set true parameter values

We define arbitrary values for the boundary separation and drift rate parameters in each experimental condition such that **each parameter is only affected by a given main effect**.

```{r}
# Only Factor A has a main effect on the boundary separation
bound <- matrix(c(3,   5.5, 
                  3.05, 5.55), nrow = L, ncol = L, byrow = TRUE)
# Only Factor B has a main effect on the drift rate
drift <- matrix(c(0.7, 0.75, 
                  2.5, 2.45), nrow = L, ncol = L, byrow = TRUE)
```

As shown in the interaction plots below, the boundary separation parameter is only affected by Factor A, while the drift rate parameter is only affected by Factor B.

```{r, echo = FALSE, fig.height = 4, fig.align = "center"}
bound_b1.color <- rgb(0.1, 0.3, 0.8, 1)  # Dark blue with transparency
bound_b2.color <- rgb(0.4, 0.8, 1.0, 1)  # Light blue with transparency
drift_b1.color <- rgb(0.5, 0.0, 0.8, 1)  # Dark purple
drift_b2.color <- rgb(0.8, 0.6, 1.0, 1)  # Light purple


layout(matrix(c(1,2), nrow = 1, ncol = 2))
par(mai = c(0.5,1,0.5,0.1), oma = c(0,0,0,0))
plot(c(0.5,1), bound[1,], type = "b", pch = 16, col = bound_b1.color, ylim = c(0, max(bound)*1.2),
     ann = FALSE,axes = FALSE, xlim = c(0.4,1.1))
axis(1, at = c(0.5,1), labels = c("Factor A1", "Factor A2"))
axis(2, at = seq(0, max(bound)*1.2,length.out = 5), labels = round(seq(0, max(bound)*1.2,length.out = 5),1), las = 2)
points(c(0.5,1), bound[2,], type = "b", pch = 16, col = bound_b2.color)
legend("bottomright", legend = c("Factor B1", "Factor B2"), col = c(bound_b1.color, bound_b2.color), pch = 16, bty = "n", cex = 1.2)
mtext(expression(bold(paste("Boundary separation ", alpha))), side = 2, line = 2.5, font = 2, cex = 1.2)

plot(c(0.5,1), drift[1,], type = "b", pch = 16, col = drift_b1.color, ylim = c(0, max(drift)*1.2),
     ann = FALSE,axes = FALSE, xlim = c(0.4,1.1))
axis(1, at = c(0.5,1), labels = c("Factor A1", "Factor A2"))
axis(2, at = seq(0, max(drift)*1.2,length.out = 5), labels = round(seq(0, max(drift)*1.2,length.out = 5),1), las = 2)
points(c(0.5,1), drift[2,], type = "b", pch = 16, col = drift_b2.color)
legend("bottomright", legend = c("Factor B1", "Factor B2"), col = c(drift_b1.color, drift_b2.color), pch = 16, bty = "n", cex = 1.2)
mtext(expression(bold(paste("Drift rate ", nu))), side = 2, line = 2.5, font = 2, cex = 1.2)
```

We prepare a list with the parameter values used to sample summary statistics for each experimental condition across experimental sessions. 

```{r}
ddm_parameters <- list(
  bound = rep(as.vector(bound), D),  # Repeated over experimental sessions
  drift = rep(as.vector(drift), D),  # Repeated over experimental sessions
  nondt = rep(0.2, D*L*L)            # Same for all conditions and experimental sessions
)
```

So that the simulation design matrix ends up looking like this:

```{r, echo = FALSE}
FA <- rep(rep(paste("A",c(1,2), sep=""), each = L), D)
FB <- rep(paste("B",c(1,2), sep=""), L*D)
Session <- rep(1:D, each = K)

design <- data.frame("Drift" = ddm_parameters$drift, 
                     "Bound" = ddm_parameters$bound, 
                     "Nondt" = ddm_parameters$nondt,
                     "Factor A" = FA, "Factor B" = FB,
                     "Condition" = rep(NA, D*K),
                     "Session" = Session)

design$Condition <- 1
design$Condition[design$Factor.B == "B2" & design$Factor.A == "A1"] <- 2
design$Condition[design$Factor.B == "B1" & design$Factor.A == "A2"] <- 3
design$Condition[design$Factor.B == "B2" & design$Factor.A == "A2"] <- 4
head(design,8)
```

### Generate EZ summary statistics

We use our custom function `simSumStats()` to generate the EZ summary statistics for each experimental session:

```{r}
set.seed(seed)
sum_stats <- simSumStats(ddm_parameters, N)
```

```{r, echo=FALSE}
full_data <- cbind(design, round(sum_stats, 4))
full_data$Factor.A <- as.factor(full_data$Factor.A)
full_data$Factor.B <- as.factor(full_data$Factor.B)
head(full_data,8)
```

## Noncrossover interaction in Mean RTs

When we plot the mean RTs simulated by condition from the selected true parameter values, we find a noncrossover interaction between Factor A and Factor B. The interaction plot shows differences in the mean RT across the two levels of both factors, with the magnitude of said difference being modulated by the second factor (e.g., the vertical lines indicating the difference between the two levels of Factor A are different for the two levels of Factor B). This is an example of a phantom interaction that can be removed by a simple scale transformation.

```{r, echo=FALSE}
layout(matrix(c(1), nrow = 1, ncol = 1))
par(mai = c(1, 1, 0.5, 0.1), oma = c(0, 0, 0, 0)) 

jitter_factor <- 0.1 
colors_b <- c("darkgreen", "chartreuse3")
diff_color <- "gray40"
names(colors_b) <- c("B1", "B2")

means <- matrix(tapply(full_data$Mrt, full_data$Condition, mean), ncol=2)

plot(NULL, xlim = c(0.5, 2.5), ylim = range(full_data$Mrt, na.rm = TRUE) * c(0.95, 1.05),      
     xaxt = "n", las = 1, ann = FALSE)     
axis(1, at = c(1, 2), labels = c("A1", "A2"), cex.axis = 1.2)
points(jitter(as.numeric(factor(full_data$Factor.A[full_data$Factor.B == "B1"])), factor = jitter_factor),
       full_data$Mrt[full_data$Factor.B == "B1"], pch = 16, col = colors_b["B1"])
points(jitter(as.numeric(factor(full_data$Factor.A[full_data$Factor.B == "B2"])), factor = jitter_factor),
       full_data$Mrt[full_data$Factor.B == "B2"], pch = 16, col = colors_b["B2"])
for(i in 1:2){
    points(c(i,i),means[,i], type="l", pch=16, cex=1.5, lty=3, lwd=4, col=diff_color)
    text(i-c(0.075,-0.075)[i], mean(means[,i]), parse(text=paste0("\"Diff\"[", i, "]")), cex = 1.2, col=diff_color)
}
for(i in 1:2){    
    points(c(1,2),means[i,], type="b", pch=16, cex=1.5, lty=2)        
}
text(0.75, max(full_data$Mrt, na.rm = TRUE)*.9, 
     expression("Diff"[1] * " " * "\u2260" * " " * "Diff"[2]), cex = 1.5, col=diff_color)
mtext("Mean Reaction Time (Mrt)", side = 2, line = 2.7, font = 2, cex = 1.4)
mtext("Factor A", side = 1, line = 2.5, font = 2, cex = 1.4)
legend("topright", legend = c("Factor B1", "Factor B2"), col = colors_b, pch = 16, bty = "n", cex = 1.5)
```


By running a frequentist ANOVA on the mean RTs generated and plotted above, we find that both Factor A and Factor B have significant main effect on the mean RTs, but more importantly, the interaction between the two factors is also statistically significant.

```{r}
# Run ANOVA test
anova_result <- aov(Mrt ~ Factor.A * Factor.B, data = full_data)
summary(anova_result)
```

These results are consistent with the Bayesian multiple regression model fit to the data below.

```{r}
modelFile <- here("output", "BUGS", "linear_model.bug")

model <- write("
model {
          # Priors
          beta0 ~ dnorm(0,1)        # Mean RT - Intercept/Baseline          
          beta1 ~ dnorm(0,1)        # Factor A - Main effect
          beta2 ~ dnorm(0,1)        # Factor B - Main effect
          beta3 ~ dnorm(0,1)        # Factor A * Factor B - Interaction
          rt_lambda ~ dunif(0,3)    # Mean RT - Precision
          
          # Condition-specific Mean RT predictions
          for(j in 1:4){
              rt_pred[j] = beta0 + beta1*A[j]+beta2*B[j]+beta3*A[j]*B[j]              
          }
          
          # Sampling model           
          for (k in 1:length(meanRT)) {
              # Person-by-condition specificDM parameters                                                  
              meanRT[k]  ~ dnorm(rt_pred[cond[k]],rt_lambda)
            }
}", modelFile)
```

```{r, warning = FALSE, results = "hide", echo=FALSE}
set.seed(seed)
# General setup
n.chains  <- 4;      n.iter    <- 3500
n.burnin  <- 250;    n.thin    <- 1

# Pass data to JAGS
data_toJAGS <- list("nTrials"  =  rep(N, length(full_data$Mrt)),
                    "meanRT"   =  full_data$Mrt,                    
                    "cond"     =  full_data$Condition,
                    "A"   =  ifelse(full_data$Factor.A == "A1", 0, 1), 
                    "B"   =  ifelse(full_data$Factor.B == "B1", 0, 1))

# Specify parameters to keep track of
parameters <- c('beta0', 'beta1', 'beta2', 'beta3')

# Prepare initial values
myinits <- rep(list(list()), n.chains)
          for(i in 1:n.chains){
              myinits[[i]] <- list(beta1 = rnorm(1,0,0.1))
          }

start <- Sys.time()
samples <- jags(data=data_toJAGS,
                parameters.to.save=parameters,
                model=modelFile,
                n.chains=n.chains,  n.iter=n.iter,
                n.burnin=n.burnin,  n.thin=n.thin,
                DIC=T, inits=myinits)
end <- Sys.time()

cat("Time taken:", round(difftime(end, start, units = "secs"), 2), "seconds")
```

```{r, echo=FALSE}
rhats <- apply(samples$BUGSoutput$sims.array,3,getRhat)
rule <- 1.05
bad.Rhat <- which(rhats>rule)
if(length(bad.Rhat) > 0){        
        which.are.bad.Rhats <- names(bad.Rhat)
        hist(rhats, breaks = 50)
        abline(v=rule, col="red", lty=2)
        legend("top",paste("Rhat > ",rule," | ",
                          (round(nrow(bad.Rhat)/(length(as.vector(rhats))),5))*100,
                          "% of chains | ", length(which.are.bad.Rhats), " chains", sep=""), lty=2, col="red", cex=0.4)
        table(which.are.bad.Rhats)
}else{      paste("No Rhat greater than ", rule, sep="")       }

# Load posterior chains
beta0 <- samples$BUGSoutput$sims.list$beta0   # Intercept
beta1 <- samples$BUGSoutput$sims.list$beta1   # Main effect of Factor A
beta2 <- samples$BUGSoutput$sims.list$beta2   # Main effect of Factor B
beta3 <- samples$BUGSoutput$sims.list$beta3   # Interaction effect
```

```{r, echo=FALSE}
layout(matrix(c(1,3,2,4), nrow = 2, ncol = 2))
par(mai = c(1, 1, 0.5, 0.1), oma = c(0, 0, 0, 0)) 

curve.col <- c("#D9FFBB", "#D9FFBB", "#CFE0C0")
breaks.col <- c("#BAF989", "#BAF989", "#AEC995")
line.col <- c("#7CC642", "#7CC642", "#7BAE4C")

xlim <- c(-3,3)
hist(beta1, main = "Factor A - Main effect", xlab = expression(beta[1]), ylab = "Frequency", col = curve.col[1], xlim = xlim, freq = FALSE, breaks = 30, border = breaks.col[1], cex.main = 1.5)
lines(density(beta1), lwd = 3, col = line.col[1])
abline(v = 0, lty = 2, col = "red")
hist(beta2, main = "Factor B - Main effect", xlab = expression(beta[2]), ylab = "Frequency", col = curve.col[2], xlim = xlim, freq = FALSE, breaks = 30, border = breaks.col[2], cex.main = 1.5)
lines(density(beta2), lwd = 3, col = line.col[2])
abline(v = 0, lty = 2, col = "red")
hist(beta3, main = "Interaction", xlab = expression(beta[3]), ylab = "Frequency", col = curve.col[3], xlim = xlim, freq = FALSE, breaks = 30, border = breaks.col[3], cex.main = 1.5)
lines(density(beta3), lwd = 3, col = line.col[3])
abline(v = 0, lty = 2, col = "red")
```

As shown in the plots above, all three posterior distributions for the regression coefficients have almost no posterior density at zero, indicating that the main and interaction effects of Factor A and Factor B on the mean RTs are supported by evidence.

## Fit a Bayesian EZ-DDM model

We now fit a Bayesian EZ-DDM model to the simulated data to properly quantify and evaluate the main and interaction effects of Factor A and Factor B on the boundary separation and drift rate parameters. <!--Here, we only show the JAGS model. The full R code can be found on the .Rmd file corresponding to this report. -->

### JAGS model

We write a JAGS model that describes the drift rate and boundary separation parameters as a function of the levels of Factor A and Factor B, incorporating main effects and the interaction between them.

\begin{equation}
\nu_{k} = \mu_{\nu} + \gamma_{1}A_{k} + \gamma_{2}B_{k} + \gamma_{3}A_{k}B_{k}\\
\end{equation}

\begin{equation}
\alpha_{k} = \mu_{\alpha} + \lambda_{1}A_{k} + \lambda_{2}B_{k} + \lambda_{3}A_{k}B_{k}
\end{equation}


```{r}
modelFile <- here("output", "BUGS", "ezddm_model.bug")

model <- write("
model {
          ##################
          ####### Priors
          ##################                    
          drift_mu ~ dnorm(0,1)                      # Baseline drift rate
          drift_lambda ~ dgamma(2,1)
          drift_sigma = pow(drift_lambda, -0.5)
          bound_mu ~ dnorm(2,1)T(0,)                 # Baseline boundary separation
          bound_lambda ~ dgamma(2,1)
          bound_sigma = pow(bound_lambda, -0.5)
          nondt ~ dexp(1)                            # Participant-specific non-decision time  
                                        
          # Regression coefficients (2 main effects + 1 interaction)
          for(i in 1:3){                           
              gamma[i] ~ dnorm(0,1)       # Drift rate regression coefficients
              lambda[i] ~ dnorm(0,1)      # Boundary separation regression coefficients
          }      
          
          # Condition-specific drift rate and boundary separation
          for(j in 1:4){
              drift_pred[j] = drift_mu + gamma[1]*A[j]+gamma[2]*B[j]+gamma[3]*A[j]*B[j]
              bound_pred[j] = bound_mu + lambda[1]*A[j]+lambda[2]*B[j]+lambda[3]*A[j]*B[j]
          }
          
          #########################
          ####### Sampling model 
          ####### (for each experimental session)
          #########################          
          for (k in 1:length(meanRT)) {
              # Person-by-condition specificDM parameters                                    
              drift[k] ~ dnorm(drift_pred[cond[k]],drift_lambda)
              bound[k] ~ dnorm(bound_pred[cond[k]],bound_lambda)                  
      
              # Forward equations from EZ Diffusion
              ey[k]  = exp(-bound[k] * drift[k])
              Pc[k]  = 1 / (1 + ey[k])
              PRT[k] = 2 * pow(drift[k], 3) / bound[k] * pow(ey[k] + 1, 2) / (2 * -bound[k] * 
                      drift[k] * ey[k] - ey[k] * ey[k] + 1)
              MDT[k] = (bound[k] / (2 * drift[k])) * (1 - ey[k]) / (1 + ey[k])
              MRT[k] = MDT[k] + nondt
      
              # Sampling distributions for summary statistics
              correct[k] ~ dbin(Pc[k], nTrials[k])
              varRT[k]   ~ dnorm(1/PRT[k], 0.5*(nTrials[k]-1) * PRT[k] * PRT[k])
              meanRT[k]  ~ dnorm(MRT[k], PRT[k] * nTrials[k])
            }
}", modelFile)
```

### Run JAGS

```{r, warning = FALSE, results = "hide"}
set.seed(seed)

# General setup
n.chains  <- 4;      n.iter    <- 3500
n.burnin  <- 250;    n.thin    <- 1

# Pass data to JAGS
data_toJAGS <- list("nTrials"  =  rep(N, length(full_data$Mrt)),
                    "meanRT"   =  full_data$Mrt,
                    "varRT"    =  full_data$Vrt,
                    "correct"  =  full_data$A,
                    "cond"     =  full_data$Condition,
                    "A"   =  ifelse(full_data$Factor.A == "A1", 0, 1), 
                    "B"   =  ifelse(full_data$Factor.B == "B1", 0, 1))

# Specify parameters to keep track of
parameters <- c('gamma', 'lambda', 'drift', 'bound', 'drift_pred', 'bound_pred')

# Prepare initial values
myinits <- rep(list(list()), n.chains)
          for(i in 1:n.chains){
              myinits[[i]] <- list(drift = rnorm(length(data_toJAGS$nTrials),1,0.1))
          }

start <- Sys.time()
samples <- jags(data=data_toJAGS,
                parameters.to.save=parameters,
                model=modelFile,
                n.chains=n.chains,  n.iter=n.iter,
                n.burnin=n.burnin,  n.thin=n.thin,
                DIC=T, inits=myinits)
end <- Sys.time()

cat("Time taken:", round(difftime(end, start, units = "secs"), 2), "seconds")

# Extract posterior samples
gamma <- samples$BUGSoutput$sims.list$gamma   # Effects on the drift rate
lambda <- samples$BUGSoutput$sims.list$lambda # Effects on the boundary separation
# Predicted drift rates and boundary separations
drift_pred <- samples$BUGSoutput$sims.list$drift_pred
bound_pred <- samples$BUGSoutput$sims.list$bound_pred
```

### Convergence diagnostics

```{r}
rhats <- apply(samples$BUGSoutput$sims.array,3,getRhat)
rule <- 1.05
bad.Rhat <- which(rhats>rule)
if(length(bad.Rhat) > 0){        
        which.are.bad.Rhats <- names(bad.Rhat)
        hist(rhats, breaks = 50)
        abline(v=rule, col="red", lty=2)
        legend("top",paste("Rhat > ",rule," | ",
                          (round(nrow(bad.Rhat)/(length(as.vector(rhats))),5))*100,
                          "% of chains | ", length(which.are.bad.Rhats), " chains", sep=""), lty=2, col="red", cex=0.4)
        table(which.are.bad.Rhats)
}else{      paste("No Rhat greater than ", rule, sep="")       }
```

### Interaction plots

We plot the posterior distributions of the drift rate and boundary separation parameters for each experimental condition in the following interaction plots. We confirm that our model is able to recover the single-main effect structure of the true parameter values used to simulate the data.

```{r, echo=FALSE, fig.width = 10, fig.height = 5}
par(mfrow = c(1,2), mar = c(5, 4, 4, 2) + 0.1)
plot_estimates(samples, "drift", colors = c(drift_b1.color, drift_b2.color))
points(c(0.9,1.9), drift[1,], type = "b", pch = 16, col = "blue")
points(c(1.1,2.1), drift[2,], type = "b", pch = 16, col = "blue")

plot_estimates(samples, "bound", colors = c(bound_b1.color, bound_b2.color))
points(c(0.9,1.9), bound[1,], type = "b", pch = 16, col = "blue")
points(c(1.1,2.1), bound[2,], type = "b", pch = 16, col = "blue")
```


### Effect sizes

Next, we look at the posterior distributions for the regression coefficients describing the main and interaction effects of Factor A and Factor B on the drift rate and boundary separation parameters. As expected, we find that Factor A only has a main effect on the boundary separation parameter, while Factor B only has a main effect on the drift rate parameter.

```{r, echo=FALSE}
# Clear any previous plot settings
par(mfrow = c(1,2), mar = c(5, 4, 4, 2) + 0.1)

# Color palette for gamma (pink family) - drift rate effects
gamma.c <- c(rgb(0.95, 0.1, 0.5, 0.4),   # Parameter 1: semi-transparent magenta pink
             rgb(0.95, 0.5, 0.7, 0.4),   # Parameter 2: semi-transparent medium pink
             rgb(1.0, 0.7, 0.9, 0.4))    # Parameter 3: semi-transparent light pink

gamma.l <- c(rgb(0.95, 0.1, 0.5, 1),     # Parameter 1: solid magenta pink
             rgb(0.95, 0.5, 0.7, 1),     # Parameter 2: solid medium pink
             rgb(1.0, 0.7, 0.9, 1))      # Parameter 3: solid light pink

# Color palette for lambda (blue family) - boundary separation effects
lambda.c <- c(rgb(0.3, 0.6, 0.9, 0.4),   # Parameter 1: semi-transparent deep blue
              rgb(0.3, 0.6, 0.9, 0.4),   # Parameter 2: semi-transparent medium blue
              rgb(0.1, 0.3, 0.8, 0.4))   # Parameter 3: semi-transparent light blue

lambda.l <- c(rgb(0.3, 0.6, 0.9, 1),     # Parameter 1: solid deep blue
              rgb(0.3, 0.6, 0.9, 1),     # Parameter 2: solid medium blue
              rgb(0.1, 0.3, 0.8, 1))     # Parameter 3: solid light blue

# Parameter names for plot titles
param_names <- c("Factor A", "Factor B", "Interaction")

# Create two side-by-side plots
for(i in 1:2) {
  # Calculate xlim for each plot based on the specific gamma and lambda pair
  xlim <- range(c(gamma[,i], lambda[,i]))
  ylim <- c(0, max(c(density(gamma[,i])$y, density(lambda[,i])$y)) * 1.1)
  
  plot(c(0,0), main = paste("Main Effect of ", param_names[i]), 
       xlab = "Effect size", ylab = "Density", 
       col = "gray", lty = 2, lwd = 1, xlim = xlim, ylim = ylim, type = "n")
  
  # Add vertical line at zero
  abline(v = 0, col = "gray", lty = 2)
  
  # Add histograms and density lines
  hist(gamma[,i], add = TRUE, freq = FALSE, breaks = 30, col = gamma.c[i])
  lines(density(gamma[,i]), lwd = 3, col = gamma.l[i])
  
  hist(lambda[,i], add = TRUE, freq = FALSE, breaks = 30, col = lambda.c[i])
  lines(density(lambda[,i]), lwd = 3, col = lambda.l[i])
  
  # Add legend to each plot
  legend("topright", 
         legend = c("Effect on drift rate", "Effect on boundary separation"), 
         col = c(gamma.l[i], lambda.l[i]), 
         lwd = 3, 
         bty = "n")
}
```




```{r, echo=FALSE}
epsilon <- 0.1
prior_constant <- pnorm(epsilon) - pnorm(-epsilon)
for(i in 1:2){
    this.gamma <- gamma[,i]
    post_mass <- mean(this.gamma > -epsilon & this.gamma < epsilon)
    this.BF <- prior_constant/post_mass
    this.BF[post_mass==0] <- Inf
    cat("The B.F. for gamma", i, " in favor of an effect is", this.BF, ".", sep="")
    if(this.BF < 1){
      cat(" Evidence in favor of the null effect!\n")
    } else if (this.BF < 1.2){
      cat(" Weak evidence in favor of the effect!\n")
    } else if (this.BF < 5){
      cat(" Moderate evidence in favor of the effect!\n")
    } else {
      cat(" Strong evidence in favor of the effect!\n")
    }
}
```


```{r, echo=FALSE}
  # Calculate xlim for each plot based on the specific gamma and lambda pair
  xlim <- range(c(gamma[,3], lambda[,3]))
  ylim <- c(0, max(c(density(gamma[,3])$y, density(lambda[,3])$y)) * 1.1)
  
  plot(c(0,0), main = "Interaction effect", 
       xlab = "Effect size", ylab = "Density", 
       col = "gray", lty = 2, lwd = 1, xlim = xlim, ylim = ylim, type = "n")
  
  # Add vertical line at zero
  abline(v = 0, col = "gray", lty = 2)
  
  # Add histograms and density lines
  hist(gamma[,3], add = TRUE, freq = FALSE, breaks = 30, col = gamma.c[3])
  lines(density(gamma[,3]), lwd = 3, col = gamma.l[3])
  
  hist(lambda[,3], add = TRUE, freq = FALSE, breaks = 30, col = lambda.c[i])
  lines(density(lambda[,3]), lwd = 3, col = lambda.l[3])
  
  # Add legend to each plot
  legend("topright", 
         legend = c("Effect on drift rate", "Effect on boundary separation"), 
         col = c(gamma.l[i], lambda.l[i]), lwd = 3, bty = "n")
```





